---
title: "Investigating Direct Estimates"
author: "Alana McGovern"
date: "`r Sys.Date()`"
output: html_document
---

```{r include=FALSE}
library(INLA)
library(SUMMER)
library(tidyverse)
library(ggpubr)
library(Rfast)
source("/Users/alanamcgovern/Desktop/Research/New Benchmarking/getAggregated.R")
inla.setOption(inla.timeout = 15)
```

In order to use the benchmarking method we have been developing, we need to have reliable direct admin1 estimates. If we don't, it is likely preferable to use a smoothed INLA model that has an admin1 fixed effect and admin2 random effect. In this report I will investigate the data requirements for adequate accuracy (MSE and coverage of CI) and precision (coefficient of variation), to try to determine guidelines for when the HT direct estimate can be used.

For each country, I used all of the information from the most recent DHS survey, but simulated the number of deaths observed from a known rate, i.e. Binomial(births observed, rate), which was different for each area but consistent across simulations.

```{r include=FALSE}
out <- NULL
countries <- c('Burundi','Ghana','Guinea','Lesotho','Liberia','Malawi','Pakistan','Rwanda','Sierra_Leone')

for(country_t in countries){
  ## load data ------
  message(paste0('Starting ',country_t,'\n'))

  data.dir <- (paste0("/Users/alanamcgovern/Desktop/Research/UN_Estimates/UN-Subnational-Estimates/Data/",country_t,'/'))
  if(file.exists(paste0(data.dir,country_t,"_cluster_dat_1frame.rda"))){
    load(paste0(data.dir,country_t,"_cluster_dat_1frame.rda"))
  }else{
    load(paste0(data.dir,country_t,"_cluster_dat.rda"))
  }
    
  dat <- mod.dat[mod.dat$age==0,]
  dat$died <- dat$Y
  dat$years.int <- as.integer(dat$years)
  dat$years <- as.numeric(as.character(dat$years))
  dat$v005 <- dat$v005/1e6

  survey_year <- max(dat$survey)
  if(country_t=='Ghana')
    survey_year <- 2014
  dat <- dat[dat$survey==survey_year,]

  # assign time periods 
  end.year <- survey_year
  if(((end.year-2000+1) %% 3)==0){
    beg.period.years <- seq(2000,end.year,3) 
    end.period.years <- beg.period.years + 2 
  }else if(((end.year-2000+1) %% 3)==1){
    beg.period.years <- c(2000,2000+2,seq(2000+4,end.year,3))
    end.period.years <- c(2000+1,2000+3,seq(2000+6,end.year,3))
  }else if(((end.year-2000+1) %% 3)==2){
    beg.period.years <- c(2000,seq(2000+2,end.year,3))
    end.period.years <- c(2000+1,seq(2000+4,end.year,3))
  }

  periods <- paste(beg.period.years, end.period.years, sep = "-") # convert the periods into string
  dat$period <- as.character(cut(dat$years, breaks = c(beg.period.years, beg.period.years[length(beg.period.years)]+5),
                               include.lowest = T, right = F, labels = periods))

  ## simulate population -----
  
  #hh.totals <- read.csv(paste0(str_remove(data.dir,paste0(country_t,'/')),'urban_frames/rwa_2012_frame_hhtotals.csv'))
  
  n_areas <- max(dat$admin1)
  dat$g_hi <- 30 # how many HH were sampled
  simdat <- dat %>% filter(period==periods[4]) %>% group_by(period,cluster,strata,urban,age,admin1.char,admin1,admin1.name) %>% 
    summarise(total = sum(total),g_hi=sum(g_hi))
  # num sampled clusters in strata
  strata_dat <- simdat %>% group_by(admin1.char,admin1.name,urban) %>% summarise(a_h = length(unique(cluster))) %>%
  # simulate number of clusters in strata
    mutate(A_h = ifelse(urban=='urban',30*a_h,40*a_h))
  # simulate number of HH in strata
  strata_dat$M_h = sapply(1:nrow(strata_dat),function(i){round(rnorm(1,150*strata_dat$A_h[i],25))})
    
  simdat <- merge(simdat,strata_dat)
  
  #define number of total births in each cluster
  simdat$N <- round(simdat$total/simdat$g_hi*rnorm(1,150,15))
  simdat$v005 <- simdat$M_h/(simdat$a_h*simdat$g_hi)/100
  
  ## run simulations ------
  
  # set parameters
  alphaU <- -3.5
  alphaR <- -3.3
  beta.adm1 <- rnorm(n_areas,0,0.1)
  d <- 0.1
  
  param_dat <- simdat %>% group_by(admin1,region = admin1.char) %>% summarise(rural = sum(N*I(urban=='rural'))/sum(N))
  param_dat$urban <- 1-param_dat$rural
  param_dat$beta <- beta.adm1
  param_dat$r <- exp(alphaU + param_dat$beta)*param_dat$urban + exp(alphaR + param_dat$beta)*param_dat$rural
  
  nsim <- 100
  coverage.dir.mat <- width.dir.mat <- deathcount.dir.mat <- error.dir.mat <- coefvar.dir.mat <- matrix(0,nsim,n_areas)
  coverage.inla1.mat <- width.inla1.mat <- deathcount.inla1.mat <- error.inla1.mat <- coefvar.inla1.mat <- matrix(0,nsim,n_areas)
  failure.inla1 <- 0
  for(k in 1:nsim){
    #if(k%%50==0)
     # cat('Starting simulation',k,'\n')
  
    # simulate total deaths in cluster 
    simdat$total_died <- sapply(1:nrow(simdat),function(i){
      rnbinom(1,(1/d)*simdat$N[i]*exp(alphaU*I(simdat$urban[i]=='urban') + alphaR*I(simdat$urban[i]=='rural') + beta.adm1[simdat$admin1[i]] + rnorm(1,0,0.01)),1/(1+d))})
    # simulate observed deaths in cluster
    simdat$died <- sapply(1:nrow(simdat),function(i){
        rhyper(1,simdat$total_died[i],simdat$N[i]-simdat$total_died[i],simdat$total[i])})

    # get direct estimate
    suppressMessages({
    dir.est <- SUMMER::getDirect(simdat, periods[4],regionVar = "admin1.char",
                                timeVar = "period", clusterVar =  "~cluster",
                                ageVar = "age", Ntrials = "total",
                                weightsVar = "v005",national.only = F)
    })
    #put in correct order
    dir.est <- dir.est[dir.est$region!='All',]
    dir.est$region_num <- unlist(lapply(str_split(dir.est$region,'_'),function(x){as.numeric(x[2])}))
    dir.est <- dir.est[order(dir.est$region_num),]
    
    # smoothed direct estimate
    
    # INLA estimates
    ## don't fit model if any area has no observed deaths
    if(sum((simdat %>% group_by(admin1.char) %>% summarise(deaths = sum(died)))$deaths==0)>0){
      inla1.est <- data.frame(admin1 = 1:n_areas,mean=NA,median=NA,lower=NA,upper=NA)
      failure.inla1 <- failure.inla1 +1
    }else{
     inla.mod1 <- try({
        INLA::inla(died ~ factor(urban) + factor(admin1) -1,
                   data=simdat, family='nbinomial', E=total,
                   control.predictor = list(compute = T, link = 1),
                   control.family = list(link = 'log'),
                   control.compute = list(config=T))})
      ## stop model if its stalling
      if(is.character(inla.mod1[1])){
         inla1.est <- data.frame(admin1 = 1:n_areas,mean=NA,median=NA,lower=NA,upper=NA)
         failure.inla1 <- failure.inla1 +1
         cat('Iteration ',k,'. Moving on.. \n')
      }else{
        inla.out <- getAggregated(inla.mod1,simdat,admin1.UR.weights = param_dat)
        inla1.est <- inla.out$adm1.est
      }
    }
    
    
    # record direct estimate info
    coverage.dir.mat[k,] <- I(dir.est$lower<param_dat$r & dir.est$upper>param_dat$r)
    deathcount.dir.mat[k,] <- (simdat %>% group_by(admin1) %>% summarise(deaths = sum(died)))$deaths
    width.dir.mat[k,] <- dir.est$upper - dir.est$lower
    coefvar.dir.mat[k,] <- (dir.est$upper - dir.est$lower)/(2*dir.est$mean)
    error.dir.mat[k,] <- dir.est$mean - param_dat$r
    
    # record INLA estimate info
    coverage.inla1.mat[k,] <- I(inla1.est$lower<param_dat$r & inla1.est$upper>param_dat$r)
    deathcount.inla1.mat[k,] <- (simdat %>% group_by(admin1) %>% summarise(deaths = sum(died)))$deaths
    width.inla1.mat[k,] <- inla1.est$upper - inla1.est$lower
    coefvar.inla1.mat[k,] <- (inla1.est$upper - inla1.est$lower)/(2*inla1.est$median)
    error.inla1.mat[k,] <- inla1.est$median - param_dat$r
  }

  area_summary <- simdat %>% group_by(region = admin1) %>% summarise(clusternum = length(unique(cluster)), obs_births = sum(total))
  area_summary$r <- param_dat$r
  area_summary$country <- country_t
  area_summary$country.num <- as.factor(which(country_t==countries))
  
  direct.summary <- inla1.summary <- area_summary
  
  direct.summary$coverage = colSums(coverage.dir.mat,na.rm=T)/(nsim-colSums(deathcount.dir.mat==0))
  direct.summary$width = colMeans(width.dir.mat,na.rm=T)
  direct.summary$mse = colMeans(error.dir.mat^2,na.rm = T)
  direct.summary$coefvar = colMeans(coefvar.dir.mat,na.rm = T)
  direct.summary$model = 'Direct'
  direct.summary$failure = colSums(deathcount.dir.mat==0)/nsim
  
  inla1.summary$coverage = colSums(coverage.inla1.mat,na.rm=T)/(nsim-colSums(deathcount.inla1.mat==0))
  inla1.summary$width = colMeans(width.inla1.mat,na.rm=T)
  inla1.summary$mse = colMeans(error.inla1.mat^2,na.rm = T)
  inla1.summary$coefvar = colMeans(coefvar.inla1.mat,na.rm = T)
  inla1.summary$model = 'INLA'
  inla1.summary$failure <- failure.inla1/nsim
  
  out <- rbind(out,direct.summary,inla1.summary)
}
```


#### How are estimates impacted by number of observed births in an area?

```{r echo=FALSE}
# how is coverage related to number of observed births?
b1 <- out %>% ggplot() + geom_point(aes(x=obs_births,y=coverage,pch=model,col=country.num)) + geom_hline(yintercept = 0.95,col='red') +
    xlab('Number of observed births') + ylab('Coverage') + scale_x_log10() +scale_shape_manual(name='Model', values = c(1,8))

# MSE?
b2 <- out %>% ggplot() + geom_point(aes(x=obs_births,y=mse,pch=model,col=country.num)) + 
    xlab('Number of observed births') +ylab('MSE') + scale_x_log10() +scale_shape_manual(name='Model', values = c(1,8))

b3 <- out %>% ggplot() + geom_point(aes(x=obs_births,y=coefvar,pch=model,group=country.num,col=country.num)) + scale_x_log10() +
    xlab('Number of observed births') + ylab('Coefficient of Variation')+scale_shape_manual(name='Model', values = c(1,8))

b1
b2
b3

b <- ggarrange(plotlist = list(b1,b2,b3),common.legend = T,legend='right')

annotate_figure(b, top = text_grob("Relationship between Admin 1 level HT estimates and number of observed births",size = 14))
```

#### How are estimates impacted by true NMR?

```{r echo=FALSE}
# how is coverage related to true rate?
r1 <- out %>% ggplot() + geom_point(aes(x=r,y=coverage,pch=model,col=country.num)) + geom_hline(yintercept = 0.95,col='red') +
    xlab('Number of observed births') + ylab('Coverage') + scale_x_log10() +scale_shape_manual(name='Model', values = c(1,8))

# MSE?
r2 <- out %>% ggplot() + geom_point(aes(x=r,y=mse,pch=model,col=country.num)) + 
    xlab('Number of observed births') +ylab('MSE') + scale_x_log10() +scale_shape_manual(name='Model', values = c(1,8))

r3 <- out %>% ggplot() + geom_point(aes(x=r,y=coefvar,pch=model,group=country.num,col=country.num)) + scale_x_log10() +
    xlab('Number of observed births') + ylab('Coefficient of Variation')+scale_shape_manual(name='Model', values = c(1,8))

rfig <- ggarrange(plotlist = list(r1,r2,r3),common.legend = T,legend='right')

annotate_figure(rfig, top = text_grob("Relationship between Admin 1 level HT estimates and true NMR",size = 14))
```



